---
title: |
  </div>
  <div style="text-align: center;">
   Y20 Placement Data Analysis Report
    </span>
    <br><span style="font-size: 24px;">MTH208 - Data Science Lab - I
    <br>GROUP-24</br>Instructor: Dr. Dootika Vats </span>
    <br><span style="font-size: 12px;">Group Members: A Atulya Sundram,Anjali Kumari, Shreyasi Kayam,Saurav Raj
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

This report provides an analysis of students placement for the batch of 2024
across all the departments of IIT Kanpur.The primary objective of this
project is to provide valuable insights on placement trends,student
performance and placement opportunity across various departments within
the college.

## About the Data

Two datasets have been used in the following:

1. Company Data
    -   **Company Name**: The name of the company offering the placement.
    -   **Profile**: The job profile or title associated with the placement.
    -   **Role**: The specific role or position offered to the student.
    -   **Location**: The location where the job or internship will take
        place.
    -   **CTC (Cost to Company)**: The total compensation offered to the
        student, usually on an annual basis, which may include salary,
        bonuses, and other benefits.
2. Student Data

-   **Student Name**: The name of the student who was placed.
-   **Roll Number**: A unique identifier for each student.
-   **Branch**: The academic branch or department of the student (e.g.,
    Computer Science, Electrical Engineering).
-   **Gender**: The gender of the student (Male, Female,NA, etc.).
-   **Blood Group**: The blood group of the student.
-   **Home State**: The state in India where the student is originally
    from.



## Obtaining the Data

-   The websites used were the following:
    - [SPO](https://spo-backend.vercel.app/): For the actual placement data
    - [P_Club](https://search.pclub.in/): For branch information and gender information of students not placed 

## Assumptions and Potential Biases

1.  **Selection Bias**: The dataset mainly contains the data for students doing their bachlor's degree (BT & BS). 
    Unfortunately, it was difficult to procure data for other branches reliably. Also, we cannot correctly determine 
    data for those students who chose to pursue higher studies.  

2.  **Accurate Role and Profile Titles**: It is assumed that the students accepted the roles and were not let off later. 

## Interesting Questions

1. Average CTC:
    - Across branches
    - Across gender
    - Across homestate

2. Placement Rate:
    - Across the same groups

3. Any biases present


```{r loading_libraries, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggthemes)
library(plotly)
library(rsconnect)
library(leaflet)
```

## Columns name

```{r conlname_of_both datasets, echo=FALSE, message=FALSE, warning=FALSE}
## Columns name 
load("companies.Rdata")
companies1 = companies2
colnames(companies1)
load("new_students.Rdata")
load("companies.Rdata")
load("new_students.Rdata")
students_ctc <- as_tibble(read.csv("students_correlated.csv"))
data_companies <- companies2
data_students <- students
data_companies <- data_companies[data_companies$company_name != "Da Vinci Derivatives B.V.", ]

colnames(students) 
```

## Branch-wise CTC Distribution

The code chunk below calculates various statistics related to the
Cost-to-Company (CTC) for each branch of students, providing insights
into placement outcomes by branch. Specifically, it summarizes:

-   **Average CTC**: The mean CTC offered to students within each
    branch.
-   **Median CTC**: The median CTC, giving a measure of central tendency
    that is less affected by extreme values.
-   **Placement Rate**: The percentage of students who received
    placement offers (calculated as the proportion of non-missing CTC
    values).
-   **Total Students**: The number of students within each branch.

These metrics allow us to compare branches in terms of both CTC offers
and placement rates, offering a more detailed view of placement
distribution across different academic branches.

```{r Branch-wise CTCs distribution, echo=FALSE, message=FALSE, warning=FALSE}
# Branch-wise CTCs distribution

overall_company_data <- data_companies
overall_student_data <- data_students


overall_student_data <- overall_student_data %>%
left_join(
    students_ctc %>% select(roll, ctc),
    by = c("roll")
)




print(n=100 ,overall_student_data %>%
   group_by(branch) %>%
   summarize(
     avg_ctc = mean(ctc, na.rm = TRUE),
     median_ctc = median(ctc, na.rm = TRUE),
     placement_rate = mean(!is.na(ctc), na.rm = TRUE) * 100,
     total_students = n(),
     .groups = "drop"
   )
)



```

## State-wise Distribution of CTCs

The code chunk below calculates key statistics on the Cost-to-Company
(CTC) based on students' home states. This analysis provides insights
into placement outcomes by state, specifically including:

-   **Average CTC**: The mean CTC offered to students from each state.
-   **Median CTC**: The median CTC, providing a robust measure of
    central tendency.
-   **Placement Rate**: The percentage of students from each state who
    received placement offers (calculated based on non-missing values
    for `company_name`).
-   **Total Students**: The total number of students from each state in
    the dataset.

These metrics offer a detailed view of how placement packages and rates
vary across different states, highlighting potential geographic trends
in placements and salary offers.

```{r state_wise, echo=FALSE, message=FALSE, warning=FALSE}
# State-wise distribution_of_CTCs

data <- overall_student_data %>%
      group_by(home_state) %>%
      summarize(
        avg_ctc = mean(ctc, na.rm = TRUE),
        median_ctc = median(ctc, na.rm = TRUE),
        placement_rate = mean(!is.na(company_name), na.rm = FALSE) * 100,
        total_students = n()
      )

print(n = 100, data)

```

## Gender-wise Placement Analysis

The following code analyzes placement data with a focus on gender,
calculating average Cost-to-Company (CTC) and placement rates across
different branches or overall, depending on the user’s selection.

-   **Data Filtering**: The data is filtered to include only students
    with specified gender information.
-   **Overall Statistics**: If the "Overall" option is selected, the
    code summarizes statistics by gender without considering specific
    branches. In this case:
    -   **Average CTC**: The mean CTC offered to students by gender.
    -   **Placement Percentage**: The percentage of students placed,
        calculated as the proportion of non-missing values in
        `company_name`.
    -   **Total Students**: The total count of students for each gender.
-   **Branch-Specific Statistics**: If "Overall" is not selected,
    branch-specific statistics are calculated by gender:
    -   **Average CTC**: The mean CTC for each gender within each
        selected branch.
    -   **Placement Percentage**: The proportion of students placed
        within each branch and gender.
    -   **Total Students**: The count of students by gender within each
        branch.

This analysis provides insights into how placement outcomes and offers
vary by gender across branches or overall, based on the user’s
preferences.

```{r gender_wise, echo=FALSE, message=FALSE, warning=FALSE}
# Gender-wise placement analysis

 data <- overall_student_data %>% filter(!is.na(gender))
    data_summary <- data %>%
    group_by(gender, branch) %>%
    summarize(
        avg_ctc = mean(ctc, na.rm = TRUE),
        placed_percent = mean(!is.na(company_name)) * 100,
        total_students = n(),
        .groups = "drop"
    ) 

print(n=100, data_summary)

```


## Plots:

Using the above compiled data, we plotted each of them in the respective tabs in the RShiny App
They can be viewed there separately


## Conclusion


Based on the analysis of the placement data for the year 2020, the
following key insights were observed:

1.  **Placement Rates**: Placement rates vary significantly across
    different branches, with certain branches having a higher placement
    rate than others. CSE has the overall highest placement rate at 80%.
    Other branches remain way behind.
    
2.  **CTC Trends**: Here again, CSE comes out on top, with an average CTC of 22 lakhs,
way above the average of 13 lakhs 

3.  **Gender Disparities**: There are noticeable **gender disparities**
    in placement rates between males and females in certain branches but, overall
    44% of females are places as compared to 41% of males. The difference can be accounted for random noise.
    Thus, we can conclude the disparity is not significant.

4.  **Regional Trends**: Students from certain **home states** are more
    likely to secure placements in high-paying roles. This may reflect
    regional demand for talent and specific recruitment patterns by
    companies in different parts of India.







